A Machine Learning Approach to MLB Catcher Framing

### 2021-12-14 Paper.Rmd
Added more to LASSO

Our methodology is to use available data to predict whether a pitch will be a strike for all data points up until the catcher catches the ball. This will give us a good idea of whether any pitch, given a number of characteristics, will be called a strike or called a ball. From the model output, we can get a probability that every pitch will be a strike and compare that to the observed outcome of the pitch. From here we need to find out which catchers during the 2021 season excelled at getting more called strikes than their probabilities say they would.

@@

@@

@@

>>>>

Our methodology is to use available data to predict whether a pitch will be a strike for all data points up until the catcher catches the ball. This will give us a good idea of whether any pitch, given a number of characteristics, will be called a strike or called a ball. From the model output, we can get a probability that every pitch will be a strike and compare that to the observed outcome of the pitch. From here we can find out which catchers during the 2021 season excelled at getting more called strikes than their probabilities say they would.

@@

The data contains pitch characteristics for every pitch thrown during the 2021 Major League Baseball season. It was scraped from [Baseball Savant](https://baseballsavant.mlb.com/) using the [baseballr](https://billpetti.github.io/baseballr/) package. The scraped data contains ~700,000 rows with ~90 columns. We are only interested in pitches that were a called strike or a ball. This is because if the batter swung at a pitch, how the catcher presents the pitch doesn't matter. This left ~350,000 rows to model from.

## Variables

@@

We chose to predict whether a pitch is a strike on several models to find the model with the highest accuracy. The model chosen will be the model used on the results.

@@

```{r}
set.seed(445)
lasso_fit <- glmnet(x[train, ], y[train], alpha = 1, lambda = lambda, family = 'binomial')
cv_lasso <- cv.glmnet(x[train, ], y[train], alpha = 1, family='binomial')

#Lambda choice
lasso_lam <- cv_lasso$lambda.min

#Get predictions
lasso_predict <- predict(lasso_fit, s = lasso_lam, newx = x[test, ], type = 'class')

#Get coefficients
lasso_coeff <- predict(lasso_fit, s = lasso_lam, type = 'coefficients')
lasso_coeff

non_zero_lasso_coef <- lasso_coeff[lasso_coeff != 0]
#Confusion matrix
lasso_mx <- table(pred=lasso_predict, true=y[test])
lasso_mx
error <- (lasso_mx[2] + lasso_mx[3])/length(lasso_predict)
1-error
```

The model obtained using the LASSO gave a similar prediction accuracy as the logistic and ridge regression models at 0.673. The LASSO also performs variable selection, and gave non-zero coefficients for:
- Pitch type
- Count
- Outs and Inning
- Batter height
- Whether the game is home or away
- Where the pitch landed
